Report on Late Winter2013 Production: Image Differencing

### 2013-03-21 W13report.tex
Added false positive summary discussion.

We next run the SourceDetectionTask on the difference image, using a
``polarity'' of {\it both} to search for both positive and negative
deviations.  We use a nominal detection threshold of 5 sigma.  In the
case of pre--filtering, we operate on the difference image directly.
In post--filtering, we correlate the difference image with its Psf (by
design, the same as the Psf of $S$) before detection is run.  All else
being equal, the images immediately preceding the detection step
($D'$) should be {\it exactly} the same.  We expect minor differences
in the variance due to the fact that the post--filtered data will have
gone through 3 convolutions (astrometric resampling; Psf--matching to
the science image; Psf filtering for detection) while the
pre--filtered data will have gone through only two (astrometric
resampling; psf--matching to the maximum likelihood image).
Differences in the shapes and sizes of the Psf--matching kernels used
in the two cases may also cause systematic differences in the images.
Detections are merged using a grow radius of 2 pixels.

Finally, we run DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement is necessarily different in the pre--filtered
and post--filtered data.  In particular, in post--filtering, detection
happens on the filtered maximum likelihood image and source
measurement on the associated pixels in the unfiltered difference
image.  In pre--filtering, the latter does not exist, and thus
detection and measurement happen on the same image.  This restricts
the types of measurements that may be made on the sources.  In
particular, the Psf flux is merely the fitted amplitude of the peak in
the image, and the Psf centroid its center.  While these measurements
are trivial to implement, it is unclear how to implement the remainder
of the measurement suite in manner comparable to the measurements on
post--filtered data.

@@

\section{Dependencies of False Positive Rate}

After finding the configurations that yielded the minimal numbers of
false positives, we then perturbed the solutions to examine how the
numbers of false positives scaled with different effects.  This
included, in order of importance for this production: the difference
image detection threshold; the number of KernelCandidates going into
the spatial model; and astrometric registration errors.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/wcs_rms.eps} 
\includegraphics[width=0.5\textwidth]{figures/wcs_offset.eps} \\
\caption{WCS}
\label{wcs_offsets}
\end{figure}

@@

\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Best Results: Production 10} \\
\hline
Visit    & Prefilter & False Positive Matches & False Positive Orphans \\
\hline
v6866601 & True      & 0                      & 70   \\
         & False     & 0                      & 143  \\
v6866602 & True      & 0                      & 42   \\
         & False     & 0                      & 53   \\
v6866603 & True      & 0                      & 23   \\
         & False     & 0                      & 36   \\
\end{tabular}
\caption{Total number of {\tt DiaSources} detected at 5--sigma from
  all 9 sensors in raft 2,2, for the configurations that yielded the
  lowest number of false positives.  Matches are determined using a
  2'' match radius with the input reference catalog.  Objects with the
  {\tt EDGE} flag are ignored. \label{tab-bestfp10}}
\end{table}

>>>>

We next run the SourceDetectionTask on the difference image, using a
``polarity'' of {\it both} to search for both positive and negative
deviations.  We use a nominal detection threshold of 5 sigma.  In the
case of pre--filtering, we detect on the difference image directly.
In post--filtering, we correlate the difference image with its Psf (by
design, the same as the Psf of $S$) before detection is run.  All else
being equal, the images immediately preceding the detection step
($D'$) should be {\it exactly} the same.  We expect minor differences
in the variance due to the fact that the post--filtered data will have
gone through 3 convolutions (astrometric resampling; Psf--matching to
the science image; Psf filtering for detection) while the
pre--filtered data will have gone through only two (astrometric
resampling; psf--matching to the maximum likelihood image).
Differences in the shapes and sizes of the Psf--matching kernels used
in the two cases may also cause systematic differences in the images.
Detections are merged using a grow radius of 2 pixels.

Finally, we run DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement is necessarily different in the pre--filtered
and post--filtered data.  In particular, in post--filtering, detection
happens on the filtered maximum likelihood image and source
measurement on the associated pixels in the unfiltered difference
image.  In pre--filtering, the latter does not exist, and thus
detection and measurement happen on the same image.  This restricts
the types of measurements that may be made on the sources.  In
particular, the Psf flux is merely the fitted amplitude of the peak in
the image, and the Psf centroid its center.  While these measurements
are trivial to implement, additional work needs to be done on how to
implement the remainder of the measurement suite in a manner comparable
to the measurements on post--filtered data.

@@

\section{False Positive Rate}

After running an ensemble of 144 pipeline configurations, we
aggregated the numbers of false detections.  We report the results for
the each of the 3 visits, and for pre-- and postfiltering for
detection.  We cut on {\tt flags.pixel.edge} to remove spurious
detections on the immediate border of the image; other than that we do
no other cuts other than requiring that the {\tt centroid.sdss} is
finite.  We count the DiaSources that match with the reference catalog
(2'' matching radius) via the {\tt refMatchId} value.  We find no
significant correlation of false detections with sensor within the
raft; therefore we report the aggregate results for all 9 sensors in
the analysis that follows.

A set of ``heat maps'' are presented in Figure~\ref{fp_heatmap}, that
demonstrate the total number of detections across all 9 sensors as a
function of the config parameters spatialKernelOrder and degGauss, for
nGauss = 3.  Qualitatively, these heat maps indicate that a Chebyshev
spatial order greater than 3 is required for all data.  In addition,
with the exception of the deconvolution case (visit 6866601 and
prefilter=False), the Gaussians need to be modified by Laguerre
polynomials of order 4 or higher.  The minimum of these detection maps
are found using the configs of highest complexity, with a very flat
slope in the number of false positives beyond spatialKernelOrder$>$3
and degGauss$>$3.  There is not much evidence for an upturn in the
numbers of false positives, meaning we have likely not reached the
case of overfitting using these configs.

The numbers of false detections in each best case are given in
Table~\ref{tab-bestfp10}.  For the best cases, there are {\it no}
false detections that associate with an input source to within 2'',
indicating that the residuals are all likely fluctuations in the
backgrounds.  We also list the ratio of negatively detected to
positively detected DiaSources.  That this ratio is significantly
greater than 1.0 in most cases suggests that there is an
oversubtraction of the background in the process, which biases this
ratio (see Section~\ref{sec-bg}).  The optimal configuartions yielding
these false detection rates are given in Table~\ref{tab-bestconfig10}.

We have examined the distribution of these false detectiosn across the
image, and find no significant overdensities near the borders of the
images, which would be expected in cases of under or over--fitting of
the various spatial models (astrometry, backgrounds, Psf--matching
kernels).  Figure~\ref{edgedist} provides the distribution of
distances from the edge of the sensor for all false detections in {\it
  blue}, with a random sample of points in the image given in {\it
  green}.  Uncertainties are plotted using the square root of the
number of points in each bin, and the histograms are normalized to
provide a probability density.  The takeaway is that we find no
significant correlation of false positives with the boundaries of the
fitting functions.

\section{False Positive Dependencies}

After finding the configurations that yielded the minimal numbers of
false positives, we then perturbed the solutions to examine how the
numbers of false positives scaled with different effects.  This
included, in order of importance for this production: the difference
image detection threshold; the number of KernelCandidates going into
the spatial model; and astrometric registration errors.  We also
undertake a theoretical analysis of the number of false positives
expected in a 4k x 4k image as a function fo detection threshold, and
examine the requirements this sets on our ability to model the
variance and backgrounds in the images.  We compare the emperical
variation of false detections with detection threshold, and show that
our data are consistent with this.

@@

%\begin{figure}
%\includegraphics[width=0.5\textwidth]{figures/wcs_rms.eps} 
%\includegraphics[width=0.5\textwidth]{figures/wcs_offset.eps} \\
%\caption{WCS}
%\label{wcs_offsets}
%\end{figure}

@@

\begin{table}
\centering
\begin{tabular}{clccc}
\hline
\multicolumn{5}{|c|}{Best Results: Production 10} \\
\hline
Visit    & Prefilter & False Detections: Matches & False Detections: Orphans & Nneg / Npos \\
\hline
v6866601 & True      & 0                         & 70  & 1.8 \\
         & False     & 0                         & 143 & 2.9 \\
v6866602 & True      & 0                         & 42  & 2.2 \\
         & False     & 0                         & 53  & 0.7 \\
v6866603 & True      & 0                         & 23  & 2.8 \\
         & False     & 0                         & 36  & 3.0 \\
\end{tabular}
\caption{Total number of {\tt DiaSources} detected at 5--sigma from
  all 9 sensors in raft 2,2, for the configurations that yielded the
  lowest number of false positives.  Matches are determined using a
  2'' match radius with the input reference catalog.  Objects with the
  {\tt EDGE} flag are ignored.  The final column lists ratios of the
  number of negatively detected false detections to those with
  positive flux. \label{tab-bestfp10}}
\end{table}
